---
title: "Exploratory Analysis"
output: html_notebook
---



# Initial Thoughts

The original article has a strong narrative and contains lots of hard data. It compares the top 50 highest paying jobs between gender in Australia, and aims to highlight the disparity between genders.

* It has many numbers, but I feel visualisations of the data may have helped convey the size of the difference, as well as more easily summarising the differences across many job roles.


# Potential Paths

* What are the job roles with the biggest disparities?
* Are there any jobs where women earn more?
  * If so, how many?
* Are there any jobs with roughly equal levels of pay (~ 10 %)
* Produce a visualisation showing the top 10 paying job roles, and the scale of the disparity between genders.


# Analysis

## Load Data

```{r}

library(readr)
library(dplyr)

jobs <- read_csv("Data/AustralianTaxableIncome.csv")


summary(jobs)
summary(jobs$X6)


tidy_jobs <- jobs %>%
            select(GenderRank = 1,
                   2:4,
                   AvgTaxIncome = 5)
                   
```
## Are there any jobs where women earn more?

```{r}
library(tidyr)
library(knitr)

Female_jobs <-tidy_jobs %>%
  filter(Gender == "Female")

Male_jobs <- tidy_jobs %>%
  filter(Gender == "Male")

wide_jobs <- inner_join(Female_jobs, Male_jobs, by = "Occupation", suffix = c("F", "M"))

  
HigherPayingFJobs <- wide_jobs %>%
  filter(AvgTaxIncomeF > AvgTaxIncomeM) %>%
  mutate(Difference = AvgTaxIncomeF - AvgTaxIncomeM) %>%
  arrange(desc(Difference))

kable(HigherPayingFJobs)

nOfFJobs <- count(HigherPayingFJobs) # Number of Jobs where women earn more
nOfJobs <- count(wide_jobs) # Number of Jobs where we have salary information for both genders.



nOfFJobs/nOfJobs * 100



```
There are `r nOfFJobs[[1]]` occupations where women earn more than men, of a total `r nOfJobs[[1]]` where we have records for both genders.

By using an inner join, we've excluded jobs where we have no records for just one gender.
* How many occupations have records for only one gender?
* Of there, what is the balance?
* Showing the difference is useful, but the difference as a percentage of the mean wage may be more insightful
* Produce a plot to show the difference between jobs?



## Job roles with the biggest disparities

```{r}
library(ggplot2)
str(wide_jobs)
wide_jobs %>%
  mutate(difference = AvgTaxIncomeM - AvgTaxIncomeF) %>%
  top_n(n = 10) %>%
  select(Occupation, AvgTaxIncomeF, AvgTaxIncomeM) %>%
  gather(key = gender, value = AvgTaxIncome, 2:3) %>%
  ggplot(aes(Occupation, AvgTaxIncome, fill = gender)) + 
    geom_bar(stat = "identity", position = "dodge") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_y_continuous(labels = scales::dollar)

```  
